---
title: "Data Analysis Exercise"
output:
  html_document:
    link-external-icon: TRUE
    toc: FALSE
bibliography: "data-analysis-exercise/data/exercise_bib.bib"
---

# Data Information

## Data Source

This data was obtained from the [Centers for Disease Control and Prevention (CDC)'s National Notifiable Diseases Surveillance System (NNDSS)](https://data.cdc.gov/NNDSS/NNDSS-Weekly-Data/x9gk-5huc).

## Data Description

Each week, the Division of Health Informatics and Surveillance at the CDC releases a weekly report describing "provisional" cases of diseases within the United States and its territories. This dataset was released on January 26, 2023 [@divisionofhealthinformaticsandsurveillance2023].

## Data Contents

This dataset contains 424,000 rows and 16 columns.

# Data Loading and Processing

## Load Libraries

```{r echo = FALSE}
# Load Libraries
library(tidyverse)
library(gtsummary)
```

## Load Data

```{r}
NNDSS_Weekly <- read_csv("data-analysis-exercise/data/NNDSS_Weekly_Data.csv")
```

## Cleaning Data

Variables of interest include state/territory, cases this week, and disease.

```{r}
# Set Cases in New York City To Be Counted as Cases in New York (State)
NNDSS_Weekly <- NNDSS_Weekly %>%
  mutate(State_Territory = case_when(
    LOCATION1 == "NEW YORK CITY" ~ "NEW YORK",
    TRUE ~ LOCATION1))

# Rename Variables
NNDSS_Weekly <- NNDSS_Weekly %>% rename(Disease = Label) %>%
  rename(Year = `Current MMWR Year`) %>% rename(Week = `MMWR WEEK`) %>% rename(Week_Cases = `Current week`)

# Remove Extra Spaces and NAs in State_Territory
NNDSS_Weekly$Disease <- sub("  ", " ", NNDSS_Weekly$Disease)
NNDSS_Weekly <- NNDSS_Weekly %>%
  filter(State_Territory != "NA")

# Preliminary Graph: Cases per State
ggplot(NNDSS_Weekly, aes(y = State_Territory, x = Week_Cases)) + geom_col()
```

# Data Summary

```{r}
tbl_summary(NNDSS_Weekly %>% select(Week_Cases, State_Territory, Disease, Year), by = Year)
```

# Save RDS File

```{r}
saveRDS(NNDSS_Weekly, file = "data-analysis-exercise/output/clean_NNDSS.rds")
```

This section was added by Hayley Hemme

Let's load the cleaned data.
```{r}
library(tidyverse)
NNDSS_clean <- readRDS("data-analysis-exercise/output/clean_NNDSS.rds")

NNDSS_clean %>% pull(Disease) %>% unique()
```

Let's look at bit more into Invasive pneumococcal disease in all ages. We'll start removing weeks with 'NA' for reported weekly cases.
```{r}
ipd <- NNDSS_clean %>% filter(Disease == "Invasive pneumococcal disease, all ages, Confirmed") %>% 
  filter(Week_Cases != "NA") 
```

Next, let's find the weekly total for 2022.
```{r message = FALSE}
weekly_total_ipd <- ipd %>%
  group_by(Week, Year) %>%
  summarise(Week_total = sum(Week_Cases))
```

Let's plot it now! Let's filter out to only include weeks in 2022 for now
```{r}
weekly_total_ipd %>% filter(Year == "2022") %>% ggplot() + stat_summary(aes(x = Week, y= Week_total), fun = "mean", geom = "line", linewidth = 1) + labs(x = "Week", y = "Cases", title = "Weekly Cases of Confirmed Invasive Pneumococcal Disease (all ages)
in the United States -- 2022") +  theme(legend.title.align = 0.5, legend.direction = "vertical", legend.box.just = "center") + theme(plot.title = element_text(hjust = 0.5, size = 12))
```
Let's also compare the cases counts from 2022 to the previous 52 weeks.
```{r message = FALSE}
weekly_total_ipd2 <- ipd %>%
  group_by(Week, Year) %>%
  summarise(Week_total_p52 = sum(`Previous 52 week Max`))

weekly_total_ipd <- full_join(ipd, weekly_total_ipd, by = c("Week", "Year"))
weekly_total_ipd <- full_join(weekly_total_ipd, weekly_total_ipd2, by = c("Week", "Year"))
```

We'll also add some design element to the plot...
```{r}
weekly_total_ipd %>% filter(Year == "2022") %>% ggplot() + stat_summary(aes(x = Week, y= Week_total), fun = "mean", geom = "line", linewidth = 1, color = "lightblue4") + labs(x = "Week", y = "Cases", title = "Weekly Cases of Confirmed Invasive Pneumococcal Disease (all ages)
in the United States -- 2021-2022", c) + stat_summary(aes(x = Week, y= `Week_total_p52`), fun = "mean", geom = "line", linewidth = 1, linetype = "dashed", color = "lightblue3") +  theme(legend.title.align = 0.5, legend.direction = "vertical", legend.box.just = "center") + theme(plot.title = element_text(hjust = 0.5, size = 12)) +     
geom_text(aes(x = 13, y = 60, label = "2022"), color = "lightblue4") + 
geom_text(aes(x = 30, y = 400, label = "2021"), color = "lightblue3")
 
```

Let's see if we can add weekly cases from 2023 so far...
```{r}
wc_2023 <-  weekly_total_ipd %>% filter(Year == "2023")
```

Now let's plot the cases from 2021 to the third week of 2023 (most recent data at time of analysis) and adjust the scale a bit.
```{r message = FALSE}
weekly_total_ipd %>% filter(Year == "2022") %>% ggplot() + stat_summary(aes(x = Week, y= Week_total), fun = "mean", geom = "line", linewidth = 1, color = "lightblue4") + labs(x = "Week", y = "Cases", title = "Weekly Cases of Confirmed Invasive Pneumococcal Disease (all ages)
in the United States -- 2021, 2022, and the first few weeks of 2023", c) + stat_summary(aes(x = Week, y= `Week_total_p52`), fun = "mean", geom = "line", linewidth = 1, linetype = "dashed", color = "lightblue3") +  theme(legend.title.align = 0.5, legend.direction = "vertical", legend.box.just = "center") + theme(plot.title = element_text(hjust = 0.5, size = 12)) +     
geom_text(aes(x = 13, y = 60, label = "2022"), color = "lightblue4") + 
geom_text(aes(x = 30, y = 400, label = "2021"), color = "lightblue3") + stat_summary(data = wc_2023, aes(x = Week, y= Week_total), geom = "line", linewidth = 1, linetype = "dotted", color = "maroon") + geom_text(aes(x = 4, y = 250, label = "2023"), color = "maroon") + scale_x_continuous(breaks = seq(1, 52, 3))
```