---
title: "Data Analysis Exercise"
output:
  html_document:
    link-external-icon: TRUE
    toc: FALSE
bibliography: "data-analysis-exercise/data/exercise_bib.bib"
---

# Data Information

## Data Source

This data was obtained from the [Centers for Disease Control and Prevention (CDC)'s National Notifiable Diseases Surveillance System (NNDSS)](https://data.cdc.gov/NNDSS/NNDSS-Weekly-Data/x9gk-5huc).

## Data Description

Each week, the Division of Health Informatics and Surveillance at the CDC releases a weekly report describing "provisional" cases of diseases within the United States and its territories. This dataset was released on January 26, 2023 [@divisionofhealthinformaticsandsurveillance2023].

## Data Contents

This dataset contains 424,000 rows and 16 columns.

# Data Loading and Processing

## Load Libraries

```{r echo = FALSE}
# Load Libraries
library(tidyverse)
library(gtsummary)
```

## Load Data

```{r}
NNDSS_Weekly <- read_csv("data-analysis-exercise/data/NNDSS_Weekly_Data.csv")
```

## Cleaning Data

Variables of interest include state/territory, cases this week, and disease.

```{r}
# Set Cases in New York City To Be Counted as Cases in New York (State)
NNDSS_Weekly <- NNDSS_Weekly %>%
  mutate(State_Territory = case_when(
    LOCATION1 == "NEW YORK CITY" ~ "NEW YORK",
    TRUE ~ LOCATION1))

# Rename Variables
NNDSS_Weekly <- NNDSS_Weekly %>% rename(Disease = Label) %>%
  rename(Year = `Current MMWR Year`) %>% rename(Week = `MMWR WEEK`) %>% rename(Week_Cases = `Current week`)

# Remove Extra Spaces and NAs in State_Territory
NNDSS_Weekly$Disease <- sub("  ", " ", NNDSS_Weekly$Disease)
NNDSS_Weekly <- NNDSS_Weekly %>%
  filter(State_Territory != "NA")

# Preliminary Graph: Cases per State
ggplot(NNDSS_Weekly, aes(y = State_Territory, x = Week_Cases)) + geom_col()
```

# Data Summary

```{r}
tbl_summary(NNDSS_Weekly %>% select(Week_Cases, State_Territory, Disease, Year), by = Year)
```

# Save RDS File

```{r}
saveRDS(NNDSS_Weekly, file = "data-analysis-exercise/output/clean_NNDSS.rds")
```
