---
title: "Tidy Tuesday Exercise"
output: 
  html_document:
    toc: FALSE
---

# Source
The GitHub repository for TidyTuesday was cloned from <https://github.com/rfordatascience/tidytuesday>, and the CSV file for 2/14/2023 was copied and pasted into the *data* folder of this portfolio.

# Load Libraries
```{r}
library(tidyverse)
```

# Load Data
```{r}
data <- read_csv("data/age_gaps.csv")
```
## Peek at Data
```{r}
str(data)
colnames(data)
```

# Data Cleaning/Manipulation

## Notes
Overall, it appears that the original data given by TidyTuesday is pretty clean already, so I will be modifying column order so that it feels easier to read across the variables. I will also be modifying gender for each character to read as "Male" or "Female" and marking each decade that the movie was released.

## Move Columns
```{r}
# Desired Order: Movie Name, Release Year, Director, 
## Actor 1 Name, Actor 1 Birthdate, Actor 1 Age, 
## Actor 2 Name, Actor 2 Birthdate, Actor 2 Age, 
## Age Difference, Couple Number

data <- data %>%
  # Actor 1
  relocate(actor_1_name, .after = director) %>%
  relocate(actor_1_birthdate, .after = actor_1_name) %>%
  relocate(actor_1_age, .after = actor_1_birthdate) %>%
  # Actor 2
  relocate(actor_2_name, .after = actor_1_age) %>%
  relocate(actor_2_birthdate, .after = actor_2_name) %>%
  relocate(actor_2_age, .after = actor_2_birthdate) %>%
  # Age Difference + Couple Number
  relocate(age_difference, .after = actor_2_age) %>%
  relocate(couple_number, .after = age_difference)

```

## Change Gender to "Male" and "Female"
```{r}
data <- data %>%
  mutate(character_1_gender = case_when(
    character_1_gender == "man" ~ "Male",
    character_1_gender == "woman" ~ "Female")) %>%
  mutate(character_2_gender = case_when(
    character_2_gender == "man" ~ "Male",
    character_2_gender == "woman" ~ "Female"))
```

## New Variable: Decade
```{r}
data <- data %>%
  mutate(Decade = case_when(
    release_year >= 1930 & release_year < 1940 ~ "1930s",
    release_year >= 1940 & release_year < 1950 ~ "1940s",
    release_year >= 1950 & release_year < 1960 ~ "1950s",
    release_year >= 1960 & release_year < 1970 ~ "1960s",
    release_year >= 1970 & release_year < 1980 ~ "1970s",
    release_year >= 1980 & release_year < 1990 ~ "1980s",
    release_year >= 1990 & release_year < 2000 ~ "1990s",
    release_year >= 2000 & release_year < 2010 ~ "2000s",
    release_year >= 2010 & release_year < 2020 ~ "2010s",
    release_year >= 2020 & release_year < 2030 ~ "2020s")) %>%
  mutate(Decade_Numeric = case_when(
    Decade == "1930s" ~ 1930, 
    Decade == "1940s" ~ 1940,
    Decade == "1950s" ~ 1950,
    Decade == "1960s" ~ 1960,
    Decade == "1970s" ~ 1970,
    Decade == "1980s" ~ 1980,
    Decade == "1990s" ~ 1990,
    Decade == "2000s" ~ 2000,
    Decade == "2010s" ~ 2010,
    Decade == "2020s" ~ 2020))
```

# Data Visualization
## Overall Look at Data
```{r}
data_summary <- do.call(cbind, lapply(data %>% select(
  Decade_Numeric, release_year, 
  actor_1_age, actor_2_age, 
  age_difference, couple_number), summary))

data_summary <- data.frame(data_summary)
rmarkdown::paged_table(data_summary)

```
Unfortunately, this feels like a poor view of the data...

## Does Age Difference Decrease by Year?
```{r}
ggplot(data, aes(x = release_year, y = age_difference)) +
  geom_point() + 
  geom_line() + 
  geom_smooth(method = "loess", color = "red", fill = "red") +
  geom_smooth(method = lm, color = "blue", fill = "blue") + 
  labs(x = "Age Difference", y = "Release Year", 
       title = "Age Difference by Release Year")
```
According to this graph, it does appear that age difference decreases as time passes. However, the lines become "messy" towards the end, so I will try to tackle this by grouping by decade.

### New Variable: Average Age Difference by Decade
```{r}
data_2 <- data %>%
  group_by(Decade) %>%
  summarise(Decade_Numeric, release_year, Decade,
            Average_Age_Difference_by_Decade =
              mean(age_difference)) %>%
  count(release_year, Decade, 
        Average_Age_Difference_by_Decade, Decade_Numeric)
```

### Graph of Average Age Difference for Each Decade
```{r}
ggplot(data = data_2, aes(x = Decade_Numeric, y = Average_Age_Difference_by_Decade)) + 
  geom_point() + 
  geom_line() + 
  geom_smooth() + 
  labs(x = "Decade", y = "Average Age Difference", 
       title = "Average Age Difference by Decade")
```
Again, it seems that there is an overall decrease in age difference from the 1940s to about the year 2000, but we see age difference increasing once again from the 2000s onward. 


## Are Men Typically Older than Women in These Movies?
```{r}
# Character 1 Age
ggplot(data, aes(x = character_1_gender, y = actor_1_age)) +
  geom_point() + 
  geom_line() + 
  geom_boxplot() + 
  labs(x = "Gender", y = "Age", 
       title = "Boxplot of Age by Gender for Character 1")

# Character 2 Age
ggplot(data, aes(x = character_2_gender, y = actor_2_age)) +
  geom_point() + 
  geom_line() + 
  geom_boxplot() + 
  labs(x = "Gender", y = "Age", 
       title = "Boxplot of Age by Gender for Character 2")

```
Yes, it does appear that men are usually slightly older than women in these films.

## Does Couple Number Impact Age Difference?
```{r}
ggplot(data, aes(x = couple_number, y = age_difference, group = couple_number)) + 
  geom_point() + 
  geom_line() + 
  geom_boxplot()
```
As we can see from the graph, the lower the couple number, the greater the age difference in general. It is interesting, though, that we see increasing age difference once again for couple numbers 6 and 7.

