---
title: "Module 8 Exercise - Fitting Basic Statistical Models: Part 2"
output: 
  html_document:
      link-external-icon: TRUE
      toc: FALSE
---

# Setup

`Tidyverse` will be used for graphing in addition to handling data processing/manipulation, and the `here` package assists in pathing. The `gtsummary` package will be used to create descriptive summary tables.

## Load Required R Packages

```{r include=FALSE}
# Data Visualization/Handling
library(tidyverse)

# Pathing
library(here)

# Generation of Descriptive Tables
library(gtsummary)
```

## Load Cleaned Data

```{r}
# Pathing
data_path <- here("fluanalysis", "data", "cleaned_data.rds")

# Data Importation
clean_data <- readRDS(data_path)
```

# Exploratory Data Analysis

To get a greater sense of the dataset's structure, I'll run the `str()` command and determine which variables of interest are numeric and/or continuous.

## Data Structure

```{r}
str(clean_data)
```

It appears that there is only one numeric/continuous variable: BodyTemp. **BodyTemp** will also serve as the main continuous outcome of interest, and **Nausea** will serve as the main categorical outcome.

Other important variables of interest (that may be associated with these outcomes) include:

-   SwollenLymphNodes
-   ChillsSweats
-   SubjectiveFever
-   Headache
-   Myalgia(Y/N)
-   EarPn
-   Vomit

### Categorical Variables: Nausea, SwollenLymphNodes, ChillsSweats, SubjectiveFever, Headache, MyalgiaYN, Myalgia, EarPn, Vomit

Since these categorical variables are just Y/N, the number of yes/no's will be presented in a table for each.

Nausea - Main Categorical Outcome
```{r}
# Table
rmarkdown::paged_table(clean_data %>% group_by(Nausea) %>% count())
```

Swollen Lymph Nodes
```{r}
rmarkdown::paged_table(clean_data %>% group_by(SwollenLymphNodes) %>% count())
```

Chills/Sweats
```{r}
rmarkdown::paged_table(clean_data %>% group_by(ChillsSweats) %>% count())
```

Subjective Fever
```{r}
rmarkdown::paged_table(clean_data %>% group_by(SubjectiveFever) %>% count())
```

Headache
```{r}
rmarkdown::paged_table(clean_data %>% group_by(Headache) %>% count())
```

Myalgia
```{r}
# Y/N
rmarkdown::paged_table(clean_data %>% group_by(MyalgiaYN) %>% count())

# None, Mild, Moderate, Severe
rmarkdown::paged_table(clean_data %>% group_by(Myalgia) %>% count())
```

Ear Pain
```{r}
rmarkdown::paged_table(clean_data %>% group_by(EarPn) %>% count())
```

Vomit
```{r}
rmarkdown::paged_table(clean_data %>% group_by(Vomit) %>% count())
```

### Continuous Variables: BodyTemp

Body Temperature - Main Continuous Outcome
```{r}
## Summary Table
Body_Temp_Summary <- do.call(cbind, lapply(
  clean_data %>% select(BodyTemp), summary))
Body_Temp_Summary <- data.frame(Body_Temp_Summary)
rmarkdown::paged_table(Body_Temp_Summary)

## Histogram
ggplot(clean_data, aes(x = BodyTemp)) + 
  geom_histogram(binwidth = 0.1) +
  labs(x = "Body Temperature", y = "Frequency", 
       title = "Frequency of Body Temperatures")
```

## Data Visualization: Predictors and Outcomes

Selected Predictors for Outcome of Nausea: Headache, EarPn, Vomit 

Selected Predictors for Outcome of Body Temperature: SwollenLymphNodes, SubjectiveFever, MyalgiaYN, Myalgia, ChillsSweats

Since it is difficult to represent categorical variables as themselves instead of numeric, I have decided to plot the **Nausea** variable as 0 for "No" and 1 for "Yes" on the y-axis.

### Ear Pain vs Nausea
```{r}
ggplot(clean_data %>% mutate(Nausea = if_else(Nausea == "Yes", 1, 0)), aes(x = EarPn, y = Nausea)) +
  geom_point() +
  geom_jitter() +
  geom_violin() +
  labs(x = "Ear Pain")
```

Most do not have ear pain. This violin plot also depicts greater lack of nausea for those who don't have ear pain. For those that do, the frequency distributions seem relatively similar, though the frequency of "Yes" to ear pain still results in slightly fewer cases of nausea.

### Vomit vs Nausea
```{r}
ggplot(clean_data %>% mutate(Nausea = if_else(Nausea == "Yes", 1, 0)), aes(x = Vomit, y = Nausea)) + 
  geom_point() + 
  geom_jitter() +
  geom_violin() +
  labs(x = "Vomiting")
```

Most do not have vomiting. However, the violin plot above shows greater frequency of nausea with vomiting than vice-versa (which makes sense).

### Headache vs Nausea
```{r}
ggplot(clean_data %>% mutate(Nausea = if_else(Nausea == "Yes", 1, 0)), aes(x = Headache, y = Nausea)) + 
  geom_point() + 
  geom_jitter() +
  geom_violin()
```

Most have a headache. From the violin graph, it appears that headache may influence nausea, though a statistical test or model should be done to verify this.

### Subjective Fever vs Body Temperature
```{r}
ggplot(clean_data, aes(x = SubjectiveFever, y = BodyTemp)) +
  geom_point() + geom_boxplot() +
  labs(x = "Subjective Fever", y = "Body Temperature")
```

Those who say "Yes" to subjective fever tend to have higher body temperatures than those who say "No".

### Swollen Lymph Nodes vs Body Temperature
```{r}
ggplot(clean_data, aes(x = SwollenLymphNodes, y = BodyTemp)) +
  geom_point() + geom_boxplot()
```

There appears to be little difference between the body temperatures of those who do have swollen lymph nodes and those who do not.

### Myalgia vs Body Temperature
```{r}
# Y/N
ggplot(clean_data, aes(x = MyalgiaYN, y = BodyTemp)) +
  geom_point() + geom_boxplot()

# Myalgia Severity
ggplot(clean_data, aes(x = Myalgia, y = BodyTemp)) +
  geom_point() + geom_boxplot()
```
It seems that there is an association between having myalgia and a higher body temperature.

### Chills/Sweats vs Body Temperature
```{r}
ggplot(clean_data, aes(x = ChillsSweats, y = BodyTemp)) +
  geom_point() + 
  geom_boxplot() +
  labs(x = "Chills/Sweats")
```
From the boxplot, it appears that those with chills/sweats tend to have higher body temperatures than those without.
