---
title: "Module 10 Exercise: Flu Analysis Modeling Workflow"
output: 
  html_document:
      toc: FALSE
---
# Flu Analysis (Logistic Regression) Model Evaluation, Performed by Kai Chen
## Load Libraries
```{r}
# Pathing
library(here)

# Data Handling
library(tidyverse)

# Model Handling and Evaluation
library(tidymodels)
```

## Load Data
```{r}
# Set Variable for Where Original Data is Located
data_location <- here("fluanalysis", "data", "cleaned_data.rds")

# Read in Original Data
original_data <- readRDS(data_location)
```

## Data Splitting
```{r}
# Set Seed for Reproducible Analysis
set.seed(2023)

# Split Original Data into Training and Testing Data
data_split <- initial_split(original_data, prop = 3/4)

# Create Data Frames from Split Data
train_data <- training(data_split)
test_data <- testing(data_split)
```

## Workflow Creation and Model Fitting
### Create Recipe for Fitting Logistic Model (Categorical Outcome)
```{r}
flu_recipe <- recipe(Nausea ~ ., data = train_data)
```

### Workflow to Create Logistic Model
```{r}
# Set Model to Logistic Regression
logistic_regression_model <- logistic_reg() %>% set_engine("glm")

# Specifying Workflow
logistic_workflow <- workflow() %>% 
  add_model(logistic_regression_model) %>%
  add_recipe(flu_recipe)

# Fitting/Training
logistic_fit <- logistic_workflow %>%
  fit(data = train_data)
```

## Model 1 Evaluation
### Prediction + ROC Curve
```{r}
# Training Data
predict(logistic_fit, train_data)
train_augment <- augment(logistic_fit, train_data)
## Generate ROC Curve
train_augment %>% 
  roc_curve(truth = Nausea, .pred_No) %>%
  autoplot()
## Calculate ROC-AUC
train_augment %>%
  roc_auc(truth = Nausea, .pred_No)


# Test Data
predict(logistic_fit, test_data)
test_augment <- augment(logistic_fit, test_data)
## Generate ROC Curve
test_augment %>%
  roc_curve(truth = Nausea, .pred_No) %>%
  autoplot()
## Calculate ROC-AUC
test_augment %>%
  roc_auc(truth = Nausea, .pred_No)
```

The fitted model appears to perform worse on the test data (ROC-AUC = 0.672) than on the training data (ROC-AUC = 0.796).

## Alternative Model (Single Predictor: RunnyNose)
### Modified Flu Recipe
```{r}
new_flu_recipe <- recipe(Nausea ~ RunnyNose, data = train_data)
```

### New Workflow
```{r}
# Specifying Workflow
new_logistic_workflow <- workflow() %>% 
  add_model(logistic_regression_model) %>%
  add_recipe(new_flu_recipe)

# Fitting/Training
new_logistic_fit <- new_logistic_workflow %>%
  fit(data = train_data)
```

### Alternative Model Evaluation
```{r}
# Training Data
predict(new_logistic_fit, train_data)
new_train_augment <- augment(new_logistic_fit, train_data)
## Generate ROC Curve
new_train_augment %>% 
  roc_curve(truth = Nausea, .pred_No) %>%
  autoplot()
## Calculate ROC-AUC
new_train_augment %>%
  roc_auc(truth = Nausea, .pred_No)


# Test Data
predict(new_logistic_fit, test_data)
new_test_augment <- augment(new_logistic_fit, test_data)
## Generate ROC Curve
new_test_augment %>%
  roc_curve(truth = Nausea, .pred_No) %>%
  autoplot()
## Calculate ROC-AUC
new_test_augment %>%
  roc_auc(truth = Nausea, .pred_No)
```

The alternative model that uses just one predictor appears to be much worse (Training ROC-AUC: 0.515, Test ROC-AUC: 0.476) than the model that uses all predictors (Training ROC-AUC: 0.796, Test ROC-AUC: 0.672).






# This section added by CONNOR H ROSS (below)

## Part II: Continous Outcome
```{r}
# Part II: Continous Outcome
## Libraries already loaded above (Thanks Kailin :))

## Set seed for reproducibility
set.seed(2)



## Split 3/4 of the data into the training set
flu_splitc1 <- initial_split(original_data, prop = 3/4)



## Create data frame for the two sets:
train_datac1 <- training(flu_splitc1)
test_datac1 <- testing(flu_splitc1)


## Full model

### Creating my recipe
flu_recipec1 <- recipe(BodyTemp ~ ., data = original_data)



### Prepare model
lin_modc1 <- linear_reg() %>%
  set_engine("lm")



### Create workflow
flu_wflowc1 <- workflow() %>%
  add_model(lin_modc1) %>%
  add_recipe(flu_recipec1)



### Prepare the recipe and train the model
flu_fitc1 <- flu_wflowc1 %>%
  fit(data = train_datac1)



### Tidy output
flu_fitc1 %>%
  extract_fit_parsnip() %>%
  tidy()



### Prediction
predict(flu_fitc1, test_datac1)



### Augment
flu_augc1 <- augment(flu_fitc1, test_datac1)



### Looks like
flu_augc1 %>%
  select(BodyTemp, .pred)



### flu_aug
flu_augc1 %>%
  rmse(BodyTemp , .pred)






### One Predictor
## Creating my recipe2
flu_recipec2 <- recipe(BodyTemp ~ RunnyNose, data = original_data)



#### Previous model will work



### Create workflow2
flu_wflowc2 <- workflow() %>%
  add_model(lin_modc1) %>%
  add_recipe(flu_recipec2)



### Prepare the recipe and train the model2
flu_fitc2 <- flu_wflowc2 %>%
  fit(data = train_datac1)



### Tidy output2
flu_fitc2 %>%
  extract_fit_parsnip() %>%
  tidy()



### Prediction2
predict(flu_fitc2, test_datac1)



### Augment2
flu_augc2 <- augment(flu_fitc2, test_datac1)



### Looks like2
flu_augc2 %>%
  select(BodyTemp, .pred)



### flu_aug2
flu_augc2 %>%
  rmse(BodyTemp , .pred)

```

# ^^ This section added by CONNOR H ROSS ^^



