---
title: "Module 8 Exercise - Fitting Basic Statistical Models: Part 1"
output: 
  html_document:
      link-external-icon: TRUE
      toc: FALSE
---

# Setup

## Load Required R Packages

Tidyverse will be used for handling data processing/manipulation, and the here() package assists in pathing.

```{r include=FALSE}
library(tidyverse)
library(here)
```

## Load Data

```{r}
# Pathing
data_path <- here("fluanalysis", "data", "SympAct_Any_Pos.Rda")

# Data Importation
raw_data <- readRDS(data_path)
```

# Data Manipulation

During this stage, unnecessary columns and any NA values will be removed according to the assignment instructions. Column removal will be accomplished through the syntax `select(!contains())`, for this method includes all columns that are not the ones specified in the command. Tidyverse has the command `drop_na()`, permitting the removal of all observations containing any NA values across the dataset.

```{r}
# Removal of Unnecessary Columns
clean_data <- raw_data %>%
  select(!contains(
    c("Activity", "Score", "FluA", "FluB", "DxName", 
      "Total", "Unique.Visit")))

rmarkdown::paged_table(clean_data)

# Remove NAs in Dataset
clean_data <- clean_data %>%
  drop_na()

rmarkdown::paged_table(clean_data)

```

# Save Cleaned Data

As specified in the assignment instructions, the resulting dataset shall be saved as an RDS file in the *data* folder of *fluanalysis*. This allows for loading of the cleaned data in another Quarto/R script.

```{r}
# Pathing
data_path <- here("fluanalysis", "data", "cleaned_data.rds")

# Saving Dataset as RDS
saveRDS(clean_data, file = data_path)
```
